---
title: "Final Project Draft"
author: "Chloe Shawah"
date: "11/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

# Load  necessary packages

library(readxl)
library(readr)
library(janitor)
library(tidyverse)
library(sf)
library(reshape2)

```

```{r Read in and Clean Scatter Plot Data}

# Clean and Prep Scatter Plot Data -----------------------------------------------------------------------------------

# Read in HDI index data, slice out miscellaneous rows, select 2017 column, rename columns

hdi <- read_xlsx("raw_data/UNDP_HDI.xlsx") %>% 
  slice(190:378) %>% 
  clean_names() %>% 
  select(country_name, x2017)
colnames(hdi) <- c("country", "hdi_val")

# Read in child mortality data, select and rename columns

child_mortality <- read_xls("raw_data/WB_Child_Mortality.xls") %>% 
  slice(4:267) %>% 
  clean_names() %>% 
  select(data_source, x63)
colnames(child_mortality) <- c("country", "cm_val")
  
# Read in GDP data, select and rename columns

gdp <- read_xls("raw_data/WB_GDP.xls") %>% 
  slice(4:267) %>% 
  clean_names() %>% 
  select(data_source, x63)
colnames(gdp) <- c("country", "gdp_val")

# Read in literacy rate data, select and rename columns

literacy <- read_xls("raw_data/WB_literacy.xls") %>% 
  slice(5:268) %>% 
  clean_names() %>% 
  select(data_source, x63)
colnames(literacy) <- c("country", "lit_val")

# Read in colonial history data, select columns

colonial_history <- read_csv("raw_data/ICOW_Colonial_History.csv") %>% 
  clean_names() %>% 
  select(state, name, col_ruler, ind_from, ind_date)

# In colonial history dataset, change ind_date variable, currently a character with a month and date - fix this by dropping the last two characters (the month) and coerce to numeric
  
colonial_history$ind_date <- substr(colonial_history$ind_date,1,nchar(colonial_history$ind_date)-2) %>% as.numeric()

# Add new column that crates factors with names for the last occupying powers based on the country codes

colonial_history <- colonial_history %>% mutate(named_ind_from = ifelse(ind_from == 200, "England", "Other")) %>% 
  mutate(named_ind_from = ifelse(ind_from == 230, "Spain", named_ind_from)) %>% 
  mutate(named_ind_from = ifelse(ind_from == 210, "Netherlands", named_ind_from)) %>% 
  mutate(named_ind_from = ifelse(ind_from == -9, "No Colonizer", named_ind_from)) %>% 
  mutate(named_ind_from = ifelse(ind_from == 220, "France", named_ind_from)) %>% 
  mutate(named_ind_from = ifelse(ind_from == 2, "United States", named_ind_from)) %>% 
  mutate(named_ind_from = ifelse(ind_from == 640, "Turkey", named_ind_from)) %>% 
  mutate(named_ind_from = ifelse(ind_from == 365, "Russia", named_ind_from)) %>% 
  mutate(named_ind_from = ifelse(ind_from == 235, "Portugal", named_ind_from))
  
# Left join the four datasets by name. This loses some countries (namely the US because of the different ways to write the name), I will work on fixing this for the final version of the project

colonial_history <- left_join(colonial_history, gdp, by = c("name" = "country"))
colonial_history <- left_join(colonial_history, child_mortality, by = c("name" = "country"))
colonial_history <- left_join(colonial_history, hdi, by = c("name" = "country"))
colonial_history <- left_join(colonial_history, literacy, by = c("name" = "country"))

# Clean the new merged dataset by rounding and coercing to numeric, save this as an object called "plot_data"

plot_data <- colonial_history %>% 
  mutate(gdp_val = round(as.numeric(gdp_val), 4)) %>% 
  mutate(hdi_val = round(hdi_val, 6)) %>% 
  mutate(cm_val = round(as.numeric(cm_val), 2))

# Write this clean data into the Shiny App directory

write.csv(plot_data, "Final_Project_Draft/plot_data.csv")


```

```{r GDP Line Plot Data}

# Clean GDP Line Plot Data ------------------------------------------------------------------------------------------

# Read in from raw-data directory, slice rows as appropriate

gdp_2 <- read_xls("raw_data/WB_GDP.xls") %>% 
  slice(3:267)
colnames(gdp_2) = gdp_2[1, ] # the first row will be the header
gdp_2 = gdp_2[-1, -2:-34]

# Gather data so each year has own observation

gdp_2 <- gdp_2 %>% 
  gather(key = year, value = gdp, "1990":"2018") 

# Round, change to numeric, and clean names

gdp_2 <- gdp_2 %>% 
  mutate(gdp = round(as.numeric(gdp), 4)) %>% 
  mutate(year = as.numeric(year)) %>% 
  clean_names()

# Join with plot_data so we can see the last occupier

gdp_line_data <- inner_join(plot_data, gdp_2, by = c("name" = "country_name"))

# Write csv file into app directory

write.csv(gdp_line_data, "Final_Project_Draft/gdp_line_data.csv")

```

```{r Child Mortality Line Plot Data}

# Clean Child Mortality Line Plot Data --------------------------------------------------------------------------------

# Read in from raw-data directory, slice rows as appropriate

cm_2 <- read_xls("raw_data/WB_Child_Mortality.xls") %>% 
  slice(3:267) 
colnames(cm_2) = cm_2[1, ]
cm_2 = cm_2[-1, -2:-4]

# Gather data so each year has own observation

cm_2 <- cm_2 %>% 
  gather(key = year, value = cm_rate, "1960":"2018")

# Round, change to numeric, and clean names

cm_2 <- cm_2 %>% 
  mutate(cm_rate = round(as.numeric(cm_rate), 2)) %>% 
  mutate(year = as.numeric(year)) %>% 
  clean_names()

# Join with plot_data so we can see the last occupier

cm_line_data <- inner_join(plot_data, cm_2, by = c("name" = "country_name"))

# Write csv file into app directory

write.csv(cm_line_data, "Final_Project_Draft/cm_line_data.csv")

```

```{r Literacy Line Plot Data}

# Clean Literacy Line Plot Data ---------------------------------------------------------------------------------------

# Read in from raw-data directory, slice rows as appropriate

lit_2 <- read_xls("raw_data/WB_literacy.xls") %>% 
  slice(4:268) 
colnames(lit_2) = lit_2[1, ]
lit_2 = lit_2[-1,-2:-14]

# Gather data so each year has own observation

lit_2 <- lit_2 %>% 
  gather(key = year, value = lit_rate, "1970":"2018") 

# Round, change to numeric, and clean names

lit_2 <- lit_2 %>% 
  mutate(lit_rate = round(as.numeric(lit_rate), 2)) %>% 
  mutate(year = as.numeric(year)) %>% 
  clean_names()

# Join with plot_data so we can see the last occupier

lit_line_data <- inner_join(plot_data, lit_2, by = c("name" = "country_name"))

# Write csv file into app directory

write.csv(lit_line_data, "Final_Project_Draft/lit_line_data.csv")

```

```{r HDI Line Plot Data}

# Clean HDI Line Plot Data ------------------------------------------------------------------------------------------

# Read in from raw-data directory, slice rows as appropriate

hdi_2 <- read_xlsx("raw_data/UNDP_HDI.xlsx") %>% 
  slice(190:378) 
hdi_2 = hdi_2[, -1:-4]
hdi_2 = hdi_2[, -30]

# Gather data so each year has own observation

hdi_2 <- hdi_2 %>% 
  gather(key = year, value = hdi, "1990":"2017") 

# Round, change to numeric, and clean names

hdi_2 <- hdi_2 %>% 
  mutate(hdi = round(as.numeric(hdi), 4)) %>% 
  mutate(year = as.numeric(year)) %>% 
  clean_names()

# Join with plot_data so we can see the last occupier

hdi_line_data <- inner_join(plot_data, hdi_2, by = c("name" = "country_name"))

# Write csv file into app directory

write.csv(hdi_line_data, "Final_Project_Draft/hdi_line_data.csv")

```


```{r Plot Drafts, include = FALSE}

# Drafts of plots ---------------------------------------------------------------------------------------------------

# For GDP plot, need to scale y with log10

gdp_plot <- ggplot(plot_data) +
  geom_point(aes(x = ind_date, y = gdp_val, color = named_ind_from)) +
  scale_y_log10()

hdi_plot <- ggplot(plot_data) +
  geom_point(aes(x = ind_date, y = hdi_val, color = named_ind_from))

cm_plot <- ggplot(plot_data) +
  geom_point(aes(x = ind_date, y = cm_val, color = named_ind_from))

lit_plot <- ggplot(plot_data) +
  geom_point(aes(x = ind_date, y = lit_val, color = named_ind_from))

# Test line plot

ggplot((gdp_line_data %>% 
          filter(named_ind_from == "Spain"))) +
  aes(x = year, y = gdp, color = name) +
  geom_line()

```

Changes to make for final version:

 - Fix the colors of the points for the occupying countries. It is confusing that the colors switch when the check boxes change.
 - I lost some points when merging by country name. There is no ISO3 code in the colonial history dataset, but perhaps I could manually add one and merge using that to eliminate lost points.
 - Include tables of data?
 - Make some regression models on scatter plots?
 - Home page with lots of disclaimers explaining issues with data: take last occupying country lightly.





